---
title: |
  Fall 2018\
  IE 451 Applied Data Analysis
author: "Taught by Savas Dayanýk"
pagetitle: IE 451 Applied Data Analysis Fall 2018
output:
  bookdown::html_document2:
    theme: yeti
    number_sections: no
    toc: yes
    toc_depth: 2
    toc_float: yes
    code_folding: show
    highlight: pygments
  word_document: default
---

-----

Assigned on Friday, October 20 2018. Due **Midnight on 22 October 2018**

**Sevil Çalýþkan**

**21701423**

-----

```{r setup, include=FALSE}
library(magrittr)
library(tidyverse)
library(knitr)
library(DT)
opts_chunk$set(echo = TRUE, collapse=TRUE)
```

# Homework 3

Answer the 8th question of **Dodgers** lab assignment (Week 04/lab) **below each part here**:

```{r table, fig.cap="The 2012 season data of Dodgers"}

d <- read_csv("dodgers.csv") %>% 
  mutate_if(is.character, factor) %>% 
  mutate(day_of_week = factor(day_of_week, levels = c("Monday", "Tuesday", "Wednesday", 
                                                      "Thursday", "Friday", "Saturday", "Sunday")), 
         month = factor(month, levels = c("APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")))

d %>% datatable(options = list(scrollX = TRUE))

```
# Questions

*Regress **attendance** on **month, day of the week**, and **bobblehead promotion**.*
```{r}

result <- d %>% lm(formula = attend~month + day_of_week + bobblehead) 

result %>% summary()
```   

## Question 1

*Is there any evidence for a relationship between attendance and other variables? Why or why not?*

Yes, some of the variables seems to have higher importance on the attendance than others. Looking to the p-values, we see that variables for month being June, day being Tuesday, Saturday or Sunday, and existing of bobblehead promotion significantly affect attendance. Other variables have higher p-values than 0.01, which causes to fail to reject the null hypothesis that $\beta_{variable} = 0$.

## Question 2

*Does the bobblehead promotion have a statistically significant effect on the attendance?*
   
Yes, bobblehead promotion seems to increase the attendance quite high (~10000), with a low p-value.   

## Question 3

*Do month and day of week variables help to explain the number of attendants?*

Some days of the week and month June seems to be effective on attendance. On Tuesdays, Saturdays and Sundays, people attend more while Friday may worth to be looked further.

## Question 4 
  
*How many fans are expected to be drawn alone by a bobblehead promotion to a home game? Give a 90\% confidence interval.*

$\beta_{bobbleheadYES}$  gives us the expected fans to be drawn alone by a bobblehead promotion. Confidence interval is calculated below. 
```{r}
coef(result)[14]
result %>% confint("bobbleheadYES", level = 0.9)
```
    
## Question 5    
    
*Predict the number of attendees to a typical home game on a Wednesday in June if a bobblehead promotion is extended. Give a 90% prediction interval.*
```{r}
predict(result, newdata= data.frame(day_of_week = "Wednesday", month= "JUN", bobblehead= "YES"), interval = "prediction", level=0.9)
```

Fit gives the prediction, lwr and upr columns are for the upper and lower limits of 90% confidence interval.
   
   <!-- * How good does the model fit to the data? Why? Comment on residual standard error and R$^2$. Plot observed attendance against predicted attendance. Print the plot as a *PDF* file. -->
  



        
